---
title: "FAOSTAT Temperature Change on Land"
output: html_notebook
---
This notebook demonstrates how to load, visualize, and forecast FAOSTAT surface temperature anomalies on land.
We will: (1) load the dataset, (2) select a country series, (3) visualize trends, and (4) fit a simple ARIMA
model for a short-horizon forecast.

Source: https://data.giss.nasa.gov/gistemp/

## Setup

```{r}
# Load required packages for data handling, modeling, and plotting
library(tspredbench)
library(daltoolbox)
library(harbinger)
library(tspredit)
```

## Load climate dataset

```{r}
# Load the climate list object and quickly inspect
data(climate)
head(climate, 5)
```

## Select a country series

```{r}
# Select Brazil's annual temperature change series
series <- climate$brazil_clima
series
```

## Visualize the series

```{r}
# Convert named vector (years as names) into numeric vectors for plotting
years <- as.numeric(names(series))
values <- as.numeric(series)

# Plot temperature anomalies for Brazil
plot(years, values, type = "l",
     main = "Surface temperature anomalies – Brazil",
     ylab = "Temperature change (°C)", xlab = "Year")
```

## Train/test split and supervised projection

```{r}
# Create time-series dataset with sliding window of 1 (one-step lag)
ts <- ts_data(values, sw = 1)

# Split into train/test (final 5 observations as test)
samp <- ts_sample(ts, test_size = 5)

# Build input/output supervised matrices
io_train <- ts_projection(samp$train)
io_test  <- ts_projection(samp$test)
```

## Fit ARIMA model

```{r}
# Create and fit ARIMA on the supervised training data
model <- ts_arima()
model <- fit(model, x = io_train$input, y = io_train$output)
```

## Forecast and evaluate

```{r}
# Forecast next 5 steps using the first test input row
prediction <- predict(model, x = io_test$input[1, ], steps_ahead = 5)

# Convert predictions and ground truth to numeric vectors
pred <- as.vector(prediction)
real <- as.vector(io_test$output)

# Evaluate predictive performance (e.g., MAE, RMSE, MAPE)
ev_test <- evaluate(model, real, pred)
ev_test
```

## References

- Hansen, J., Ruedy, R., Sato, M., & Lo, K. (2010). Global surface temperature change. Reviews of Geophysics, 48(4).
  https://doi.org/10.1029/2010RG000345
