---
title: "M1 Competition Time Series"
output: html_notebook
---
This notebook demonstrates loading and visualizing time series from the first Makridakis forecasting competition (M1).
We select a monthly subset and build a simple ARIMA forecast on a single series.

Source: https://onlinelibrary.wiley.com/doi/10.1002/for.3980010202

## Setup

```{r}
# Load required packages
library(tspredbench)
library(daltoolbox)
library(harbinger)
library(tspredit)
```

## Load M1 dataset

```{r}
# Load the consolidated M1 object (list with frequency buckets)
data(m1)
head(m1$monthly, 5)
```

## Select and prepare a series

```{r}
# Select the monthly data frame and extract a single row as a numeric series
df <- m1$monthly
head(df, 5)

values <- as.numeric(df[2, ])           # choose the second series for illustration
values <- na.omit(values)                # remove missing values if any
points <- length(values)
time <- 1:points
```

## Visualize the series

```{r}
plot(time, values, type = "l",
     main = "M1 monthly series",
     ylab = "Value", xlab = "Time index")
```

## Train/test split and supervised projection

```{r}
ts <- ts_data(values, sw = 1)
samp <- ts_sample(ts, test_size = 5)
io_train <- ts_projection(samp$train)
io_test  <- ts_projection(samp$test)
```

## Fit ARIMA model

```{r}
model <- ts_arima()
model <- fit(model, x = io_train$input, y = io_train$output)
```

## Forecast and evaluate

```{r}
prediction <- predict(model, x = io_test$input[1, ], steps_ahead = 5)
pred <- as.vector(prediction)
real <- as.vector(io_test$output)
ev_test <- evaluate(model, real, pred)
ev_test
```

## References

- Makridakis, S., Wheelwright, S. C., & Hyndman, R. J. (1998). Forecasting: Methods and Applications.
- Makridakis, S., et al. (1982). The accuracy of extrapolation (time series) methods: Results of a forecasting competition. Journal of Forecasting.
