---
title: "M3 Competition Time Series"
output: html_notebook
---
This notebook demonstrates loading and visualizing time series from the third Makridakis forecasting competition (M3).
We select a monthly series and build a simple ARIMA forecast.

Source: https://www.sciencedirect.com/science/article/abs/pii/S0169207000000571

## Setup

```{r}
# Load required packages
library(tspredbench)
library(daltoolbox)
library(harbinger)
library(tspredit)
```

## Load M3 dataset

```{r}
data(m3)
head(m3$monthly, 5)
```

## Select and prepare a series

```{r}
series <- m3$monthly$M1111
values <- as.numeric(series)
time <- seq_along(values)
```

## Visualize the series

```{r}
plot(time, values, type = "l",
     main = "M3 monthly series: M1111",
     ylab = "Value", xlab = "Time index")
```

## Train/test split and supervised projection

```{r}
ts <- ts_data(values, sw = 1)
samp <- ts_sample(ts, test_size = 5)
io_train <- ts_projection(samp$train)
io_test  <- ts_projection(samp$test)
```

## Fit ARIMA model

```{r}
model <- ts_arima()
model <- fit(model, x = io_train$input, y = io_train$output)
```

## Forecast and evaluate

```{r}
prediction <- predict(model, x = io_test$input[1, ], steps_ahead = 5)
pred <- as.vector(prediction)
real <- as.vector(io_test$output)
ev_test <- evaluate(model, real, pred)
ev_test
```

## References

- Makridakis, S., & Hibon, M. (2000). The M3-Competition: results, conclusions and implications. International Journal of Forecasting.
