---
title: "M4 Competition Time Series"
output: html_notebook
---
This notebook demonstrates loading and visualizing time series from the fourth Makridakis forecasting competition (M4).
We select a monthly series and build a simple ARIMA forecast.

Source: https://github.com/Mcompetitions/M4-methods

## Setup

```{r}
# Load required packages
library(tspredbench)
library(daltoolbox)
library(harbinger)
library(tspredit)
```

## Load M4 dataset

```{r}
data(m4)
head(m4$monthly$M32000, 5)
```

## Select and prepare a series

```{r}
series <- m4$monthly$M32000
values <- as.numeric(series)
values <- na.omit(values)
time <- seq_along(values)
```

## Visualize the series

```{r}
plot(time, values, type = "l",
     main = "M4 monthly series: M32000",
     ylab = "Value", xlab = "Time index")
```

## Train/test split and supervised projection

```{r}
ts <- ts_data(values, sw = 1)
samp <- ts_sample(ts, test_size = 5)
io_train <- ts_projection(samp$train)
io_test  <- ts_projection(samp$test)
```

## Fit ARIMA model

```{r}
model <- ts_arima()
model <- fit(model, x = io_train$input, y = io_train$output)
```

## Forecast and evaluate

```{r}
prediction <- predict(model, x = io_test$input[1, ], steps_ahead = 5)
pred <- as.vector(prediction)
real <- as.vector(io_test$output)
ev_test <- evaluate(model, real, pred)
ev_test
```

## References

- Makridakis, S., Spiliotis, E., & Assimakopoulos, V. (2020). The M4 Competition: Results, findings, conclusion and way forward. International Journal of Forecasting.
