---
title: "Pesticides Use Statistics"
output: html_notebook
---
This notebook demonstrates loading, visualizing, and forecasting FAOSTAT pesticides use series for a selected country
using a simple ARIMA model.

Source: http://www.fao.org/faostat/en/#data/RP

## Setup

```{r}
# Load required packages
library(tspredbench)
library(daltoolbox)
library(harbinger)
library(tspredit)
```

## Load pesticides dataset

```{r}
data(pesticides)
head(pesticides, 5)
```

## Select a country series

```{r}
series <- pesticides$brazil_pest
series
```

## Visualize the series

```{r}
years <- as.numeric(names(series))
values <- as.numeric(series)
plot(years, values, type = "l",
     main = "Pesticides use – Brazil",
     ylab = "Tonnes", xlab = "Year")
```

## Train/test split and supervised projection

```{r}
ts <- ts_data(values, sw = 1)
samp <- ts_sample(ts, test_size = 5)
io_train <- ts_projection(samp$train)
io_test  <- ts_projection(samp$test)
```

## Fit ARIMA model

```{r}
model <- ts_arima()
model <- fit(model, x = io_train$input, y = io_train$output)
```

## Forecast and evaluate

```{r}
prediction <- predict(model, x = io_test$input[1, ], steps_ahead = 5)
pred <- as.vector(prediction)
real <- as.vector(io_test$output)
ev_test <- evaluate(model, real, pred)
ev_test
```

## References

- FAOSTAT Pesticides Use – Domain information: RP_e_README_Domain_Information_2024.
