---
title: "IBOVESPA's 50 Most Traded Stocks"
output: html_notebook
---
This notebook demonstrates how to load, visualize, and forecast closing prices for a selected IBX‑50 ticker
from the `stocks` dataset using a simple ARIMA model.

Source: https://www.b3.com.br

## Setup

```{r}
# Load required packages
library(tspredbench)
library(daltoolbox)
library(harbinger)
library(tspredit)
```

## Load stocks dataset

```{r}
# Load the stocks list object and inspect a known ticker (if present)
data(stocks)
if ("VALE3" %in% names(stocks)) head(stocks$VALE3, 5) else head(stocks[[1]], 5)
```

## Select a ticker series

```{r}
# Prefer VALE3 if available, otherwise pick the first ticker
if ("VALE3" %in% names(stocks)) {
  series_df <- stocks$VALE3
  ticker <- "VALE3"
} else {
  ticker <- names(stocks)[1]
  series_df <- stocks[[1]]
}
head(series_df, 5)
```

## Visualize the closing price

```{r}
years <- series_df$date
values <- series_df$close
plot(years, values, type = "l",
     main = paste0("Closing Price (", ticker, ")"),
     ylab = "Close (R$)", xlab = "Date")
```

## Train/test split and supervised projection

```{r}
ts <- ts_data(values, sw = 1)
samp <- ts_sample(ts, test_size = 5)
io_train <- ts_projection(samp$train)
io_test  <- ts_projection(samp$test)
```

## Fit ARIMA model

```{r}
model <- ts_arima()
model <- fit(model, x = io_train$input, y = io_train$output)
```

## Forecast and evaluate

```{r}
prediction <- predict(model, x = io_test$input[1, ], steps_ahead = 5)
pred <- as.vector(prediction)
real <- as.vector(io_test$output)
ev_test <- evaluate(model, real, pred)
ev_test
```

## References

- Box, G. E. P., Jenkins, G. M., Reinsel, G. C., & Ljung, G. M. (2015). Time Series Analysis: Forecasting and Control.
