---
title: "Gecco Challenge Dataset - Water Quality"
output: html_notebook
---
Data collection for water quality monitoring

* Multivariate series with labeled anomalies.
* Recommended use: multivariate or univariate event detection

Source: https://www.spotseven.de/gecco/gecco-challenge


## Load series
```{r}
library(dalevents)
library(daltoolbox)
library(harbinger)


## Load series ----------------------
data(gecco)
plot(as.ts(gecco$gecco[,1:9]))
```

Gecco recommended sample: One day with anomalie

```{r}
data <- gecco$gecco
data <- data[16500:18000,]
plot(as.ts(data))
```

## Preprocessing

* Normalize data

```{r}
ts <- data[,1:9]

preproc <- ts_norm_gminmax()
preproc <- fit(preproc, ts)
ts <- transform(preproc, ts)

head(ts)
```
```{r}
plot(as.ts(ts))
```


## Univariate series selection
Select the desired variable directly from preprocessed data.

```{r}
series <- ts$ph
plot(as.ts(series))
```

## Event detection experiment

Creating a data frame to organize experiment results.

```{r}
#Experiments results organization
experiment <- data.frame(method="hanr_arima",
                         dataset="Gecco",
                         series="ph",
                         elapsed_time_fit=0,
                         elapsed_time_detection=0,
                         accuracy=0,
                         precision=0,
                         recall=0,
                         F1=0)

head(experiment)
```
Detection steps
```{r}
#Establishing arima method
model <- hanr_arima()
```


```{r}
#Fitting the model
s <- Sys.time()
model <- fit(model, series)
t_fit <- Sys.time()-s
```


```{r}
#Making detections
s <- Sys.time()
detection <- detect(model, series)
t_det <- Sys.time()-s
```


## Results analysis


```{r}
#Filtering detected events
print(detection |> dplyr::filter(event==TRUE))
```

Visual analysis
```{r}
#Ploting the results
grf <- har_plot(model, series, detection, data$event)
plot(grf)
```

Evaluate metrics
```{r}
#Evaluating the detection metrics
ev <- evaluate(model, detection$event, data$event)
print(ev$confMatrix)
```

Recording experiment results
```{r}
#Experiment update
#Time
experiment$elapsed_time_fit[1] <- as.numeric(t_fit)*60
experiment$elapsed_time_detection[1] <- as.numeric(t_det)*60

#Metrics
experiment$accuracy[1] <- ev$accuracy
experiment$precision[1] <- ev$precision
experiment$recall[1] <- ev$recall
experiment$F1[1] <- ev$F1

print(experiment)
```

### Add new series to experiment 
Repeat detection steps

* Add new rows to the `experiment` data frame
* Run detection steps (create model, fit and detect) for the new series
* Update `experiment` with the new series result
  * Repeated steps for didactic purposes
  * WARNING: In real experimental situations, variable selection and repetition of detection steps should be encapsulated in a loop or function

### Experiment record
The retain experiments data record results

* Save detection results: Save `detection` object after detection of each series
* Save experiment: Save `experiment` object after finishing the complete experiment
